Class-based network segregation, Economic Inequality and Redistributive Preferences across societies

### 2024-03-29 paper1_jiturra.Rmd
ajustes tabla descriptiva y resultados macro

texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1),
       caption = paste("(\\#tab:table1)","Multilevel models for Class-based network segregation and Redistributive Preferences"),
       custom.note = "%stars. Models include sampling weights. Gender, age, marital status and religion are included as controls",
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,
       symbol = "+",scalebox = 0.60,
      groups = list("Social Class (Ref.= Service Class)" = 4:5, "Household Income (Ref.= Tertile I)"= 7:9,  "Partner's Social Class (Ref.= Service Class)" = 12:14, "Homogeneity x Social Class"=15:16),
      custom.gof.rows = list("Controls"=rep("Yes",6)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"),
       float.pos = "h!",
       use.packages = F)
```

@@

Model 1 integrates our measure of network homogeneity. Initially, as previously noted in the macro associations, individuals with more homogeneous networks tend to exhibit stronger redistributive preferences. This association persists in Model 2, where we control for network size and social trust—two pivotal factors in establishing and consolidating interpersonal networks. However, after including social class into the analysis, the relationship between homogeneity and redistributive preferences appears to reverse, with a reduction in statistical significance (p<0.10).

@@

texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.60,
       caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences",
       custom.note = "%stars. Models include sampling weights and individual level controls centered within cluster (group mean)",
       caption.above = T,leading.zero = T,single.row = T,
       custom.coef.map = ccoef,
      groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11),
       float.pos = "h!",
      custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups",
                           "Var: Group (Intercept)",
                           "Var: Group Homogeneity",
                           "Var: Group Intermediate Class ",
                           "Var: Group Working Class",
                           "Cov: Group (Intercept), Homogeneity",
                           "Cov: Group (Intercept), Intermediate Class ",
                           "Cov: Group (Intercept), Working Class",
                           "Cov: Group Homogeneity, Intermediate Class",
                           "Cov: Group Homogeneity, Working Class",
                           "Cov: Group Intermediate Class, Working Class",
                           "Var: Residual"),
       use.packages = F)
```

```{r fig5, fig.dim=c(12,7),fig.cap='Three-way interaccion effects for Redistributive Preferences, Network segregation, Social class and Income Inequality'}
### Predicted values homo*class*GiniD ---------------------------------------
# ginid_max<- round(max(dfreg$gini_disp) ,2)
# ginid_min<- round(min(dfreg$gini_disp),2)

@@

[Here Descriptive table]

@@

texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.60,
       caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences",
       custom.note = "%stars. Models include sampling weights and individual level controls centered within cluster (group mean)",
       caption.above = T,leading.zero = T,single.row = T,
       custom.coef.map = ccoef,
      groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11),
       float.pos = "h!",
      custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups",
                           "Var: Group (Intercept)",
                           "Var: Group Homogeneity",
                           "Var: Group Intermediate Class ",
                           "Var: Group Working Class",
                           "Cov: Group (Intercept), Homogeneity",
                           "Cov: Group (Intercept), Intermediate Class ",
                           "Cov: Group (Intercept), Working Class",
                           "Cov: Group Homogeneity, Intermediate Class",
                           "Cov: Group Homogeneity, Working Class",
                           "Cov: Group Intermediate Class, Working Class",
                           "Var: Residual"),
       use.packages = F)
```

@@

texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1),
       caption = paste("(\\#tab:table2)","Multilevel models for Class-based network segregation and Redistributive Preferences"),
       custom.note = "%stars. Models include sampling weights. Gender, age, marital status and religion are included as controls",
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,
       symbol = "+",scalebox = 0.60,
      groups = list("Social Class (Ref.= Service Class)" = 4:5, "Household Income (Ref.= Tertile I)"= 7:9,  "Partner's Social Class (Ref.= Service Class)" = 12:14, "Homogeneity x Social Class"=15:16),
      custom.gof.rows = list("Controls"=rep("Yes",6)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"),
       float.pos = "h!",
       use.packages = F)
```

>>>>

texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1),
       caption = paste("(\\#tab:table1)","Multilevel models for Class-based network segregation and Redistributive Preferences"),
       custom.note = "Note: Models include sampling weights. Gender, age, marital status and religion are included as controls. Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,
       symbol = "+",scalebox = 0.60,
      groups = list("Social Class (Ref.= Service Class)" = 4:5, "Household Income (Ref.= Tertile I)"= 7:9,  "Partner's Social Class (Ref.= Service Class)" = 12:14, "Homogeneity x Social Class"=15:16),
      custom.gof.rows = list("Controls"=rep("Yes",6)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"),
       float.pos = "h!",
       use.packages = F)
```

@@

Model 1 integrates our measure of network homogeneity. Initially, as previously noted in the macro associations, individuals with more homogeneous networks tend to exhibit stronger redistributive preferences. This association persists in Model 2, where we control for network size and social trust—two pivotal factors in establishing and consolidating interpersonal networks. However, after including social class into the analysis, the relationship between homogeneity and redistributive preferences appears to reverse, with a reduction in statistical significance (_p_<0.10).

@@

texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.7,
       caption = paste("(\\#tab:table2)","Multilevel models for Income Inequality, Network segregation and Redistributive Preferences"),
       custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,single.row = T,
       custom.coef.map = ccoef,
      groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11),
       float.pos = "h!",
      custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups",
                           "Var: Group (Intercept)",
                           "Var: Group Homogeneity",
                           "Var: Group Intermediate Class ",
                           "Var: Group Working Class",
                           "Cov: Group (Intercept), Homogeneity",
                           "Cov: Group (Intercept), Intermediate Class ",
                           "Cov: Group (Intercept), Working Class",
                           "Cov: Group Homogeneity, Intermediate Class",
                           "Cov: Group Homogeneity, Working Class",
                           "Cov: Group Intermediate Class, Working Class",
                           "Var: Residual"),
       use.packages = F)
```

Table \@ref(tab:table2) presents the results of the multilevel models for income inequality. In Model 1, income inequality demonstrates a positive association, as expected, yet it fails to reach significance. Subsequently, Model 2 incorporates economic prosperity through GDP per capita, revealing a negative and significant relationship (p > .05). Furthermore, Model 3 indicates that including the size of the welfare state establishes a positive and significant relationship (_p_ > .05) while GDP per capita maintains its negative association with redistributive preferences (_p_ > .01). In other words, in Model 3, economic prosperity appears to depress redistributive preferences; however, when contextualized within a larger welfare state, greater redistributive preferences are observed.

Finally, Model 4 incorporates the three-way interaction to determine whether the relationship established in $H_1$ is moderated by income inequality. The relationship observed at the micro level retains its original meaning, indicating that segregation in terms of network homogeneity polarizes the economic interests of the classes. Additionally, when observing how this relationship is moderated by the level of inequality in the country, it appears to mitigate, with its intensity decreasing as inequality increases. Figure \@ref(fig:fig5) illustrates how the relationship between network homogeneity and social class is more pronounced when inequality is low and gradually diminishes until it becomes almost absent when inequality is high.

Overall, while Hypothesis 2a posited that greater income differences would amplify social segregation and, consequently, polarize redistributive preferences to a greater extent, the results provide evidence in favor of Hypothesis 2b. This suggests that inequality plays a mitigating role in class segregation regarding redistributive preferences.

```{r fig5, fig.dim=c(12,7),fig.cap='Three-way interaccion Income Inequality, Network segregation and  Redistributive Preferences'}
### Predicted values homo*class*GiniD ---------------------------------------
# ginid_max<- round(max(dfreg$gini_disp) ,2)
# ginid_min<- round(min(dfreg$gini_disp),2)

@@

```{r descriptive, results='asis', cache=TRUE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,sjlabelled,haven,vtable)  
rm(list=ls())

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  religion,
  partner,
  union,
  workst,
  "gini_disp"=wid_gini_disp,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  gdppercapita,
  country2,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         gini_disp=scale(gini_disp))

dfreg <- dfreg %>% dplyr::select(
  egal,
  homclass3_V_res,
  know_total,
  socialtrust,
  dclass3res_V,
  Q03pcm,
  edyears,
  dclass3spo_V,  
  workst,
  union,
  female,
  agenum,
  religion,
  partner,
  "gini_disp",,
  loggdppercapita,
  wstate2
  )

varlab <- 
c("Redistributive preferences","Class-based network homogeneity",
  "Network size","Social trust","Social class",
  "Household Income","Education in years",
  "Partner's Social Class","Labor status","Union membership",
  "Gender","Age in years","Religion","Has partner",
  "Income Inequality - Gini Index (z-score)",
  "GDP/capita (z-score)",
  "Size of the welfare state(z-score)")

sjlabelled::set_label(dfreg) <- varlab
# names(dfreg) <- varlab

st(dfreg,labels=T,
   title="Descriptive Statistics for Study Variables",
   out = "latex",
   digits = 1, 
   factor.counts = F,
   factor.numeric = F,
   summ = list(
     c('notNA(x)','mean(x)','sd(x)','min(x)','max(x)'),
     c('notNA(x)','mean(x)')
   ),
   summ.names = list(
     c('N','Mean','SD','Min','Max'),
     c('Count','Percent')   
   ))

@@

texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.60,
       caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences",
       custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,single.row = T,
       custom.coef.map = ccoef,
      groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11),
       float.pos = "h!",
      custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups",
                           "Var: Group (Intercept)",
                           "Var: Group Homogeneity",
                           "Var: Group Intermediate Class ",
                           "Var: Group Working Class",
                           "Cov: Group (Intercept), Homogeneity",
                           "Cov: Group (Intercept), Intermediate Class ",
                           "Cov: Group (Intercept), Working Class",
                           "Cov: Group Homogeneity, Intermediate Class",
                           "Cov: Group Homogeneity, Working Class",
                           "Cov: Group Intermediate Class, Working Class",
                           "Var: Residual"),
       use.packages = F)
```

@@

texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1),
       caption = paste("(\\#tab:table2)","Multilevel models for Class-based network segregation and Redistributive Preferences"),
       custom.note = "Note: Models include sampling weights. Gender, age, marital status and religion are included as controls. Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,
       symbol = "+",scalebox = 0.60,
      groups = list("Social Class (Ref.= Service Class)" = 4:5, "Household Income (Ref.= Tertile I)"= 7:9,  "Partner's Social Class (Ref.= Service Class)" = 12:14, "Homogeneity x Social Class"=15:16),
      custom.gof.rows = list("Controls"=rep("Yes",6)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"),
       float.pos = "h!",
       use.packages = F)
```
